### 2021-06-05 Paper_Outline.Rmd

---
title: "Paper Outline"
author: "Haroon Atcha"
date: "6/4/2021"
output: html_document
---

# Summary

Outlines help me write. They are good at organizing my thoughts and help me put citations and figures in the right place. A clean outline makes draft writing much easier. This is the outline for the paper that should hopefully come out of this project.

# Outline

1. Introduction
    - Problems
      - Community Colleges operate under unique data constraints, most of the literature on enrollment management is for selective-enrollment universities
        - Cite: Aksenova, Chen, Nandeshwar, Slim (From selective enrollment universities)
        - cite: Bender, Lawrence, Pennington (from community colleges)
      - Predictive models of enrollment don't reflect our theoretical models
        - Not all processes are ARIMA; cite: Chen, Pfitzner
        - You should model different processes differently; cite: Chen, Trusheim
    - Argument: We can do better by having our statistical models match our theoretical models

2. Data
    - Show Data Generation Chart
    - Data Generation Process
      - Returning students (Linear model)
      - New students (Linear with time series terms)
    - Summary Statistics
      - Prove that the summary statistics are close to what we'd expect in an actual case study
      - Make the case for parsimonious models (since I only used 1-2 variables for each process)
      - Show enrollment time series

3. Methods
    - Aggregate models
      - ARIMA
      - Linear model
    - Stacked model
      - Logit component (returning)
      - ARIMA component (new enrollees)

4. Results
    - Robustness checks
      - Individual level confusion matrix
      - MAPE, MAE, and MSE table for each model type
        - ARIMA only
        - Linear only
        - Stacked linear
        - Stacked curvilinear
    - Toy example has prediction rates close to what case studies show
    - **Even a misspecified stacked model predicts better if it's closer to the theoretical process**

5. Conclusion
    - Predictive models perform better when they are reflective of our theoretical knowledge